---
title: "Getting Started with JW26ADS8192"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ADS8192}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

JW26ADS8192 makes it easy to run Differential Gene Expression Analysis (via DESeq2) using RNA-seq data stored as sample metadata an counts matrices (TSV/CSV).

## Quick Example


### Step 1: Install JW26ADS8192 from GitHub
```{r JW26ADS8192_installation}
remotes::install_github("wilsonjewel27/JW26ADS8192")
```

### Step 2: Load the following libraries &example dataset
```{r library_import}
library(JW26ADS8192)
library(ggplot2)
library(SummarizedExperiment)
library(DESeq2)
library(apeglm)    
library(stats)
library(rlang)
library(utils)
library(readr)

ex_counts <- readr::read_tsv("./JW26ADS8192/tests/testdata/example_counts.tsv")
ex_meta   <- readr::read_tsv("./JW26ADS8192/tests/testdata/example_meta.tsv")
```

### Start the Differential Gene Expressional Analysis

#### Step 1: Determine the Low Expression Filter Threshold
Evaluate model performance across different threshold values and select the best one. 
```{r determine_filter_threshold}
JW26ADS8192 determine_filter_threshold --count ex_counts --meta ex_meta --out results/
```

# Step 2; Filter low expression genes
se_filtered <- filter_low_exp_genes(example_se, min_count_per_group = 10)

# Step 3: Run the DESeq2 pipeline to get differential gene expression results
se_dge<- run_DESeq2(se_filtered)

# Step 4: Run the log2_shrinkage function on the results of the DESeq2 function to create more reliable estimates
se_dge_shrink <- log2_shrinkage(se_dge)

# Step 5: Create a volcano plot of the gene expression using the results obtained from the log2_shrinkage function
example_se_volcano<- generate_volcano(se_dge_shrink)

# Step 6: Generate the regulation summary of the genes
DESeq2_gene_reg_summary<- gene_regulation_summary(se_dge_shrink)

# Step 7: Export final results
example_se_exports<- export_outputs(res_df = se_dge_shrink, summary_df = DESeq2_gene_reg_summary, filtering_diag = example_se_filtering_assessment, volcano = example_se_volcano, output_dir = file.path(tempdir(), "de_output") )
```

