---
title: "Getting Started with JW26ADS8192"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ADS8192}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

JW26ADS8192: Streamlined Differential Gene Expression Analysis with DESeq2 
Designed for RNA-seq workflows, this package provides a streamlined pipeline from raw count matrices and sample metadata (TSV/CSV) to differential expression results.

## Quick Example

### Step 1: Install JW26ADS8192 from GitHub
```{r JW26ADS8192_installation}
remotes::install_github("wilsonjewel27/JW26ADS8192")
```

### Step 2: Load the following libraries &example dataset
```{r library_import}
library(JW26ADS8192)
library(ggplot2)
library(SummarizedExperiment)
library(DESeq2)
library(apeglm)    
library(stats)
library(rlang)
library(utils)
library(readr)

ex_counts <- readr::read_tsv("./JW26ADS8192/tests/testdata/example_counts.tsv")
ex_meta   <- readr::read_tsv("./JW26ADS8192/tests/testdata/example_meta.tsv")
```

### Step 3: Start the Differential Gene Expressional Analysis

#### Step 1: Determine the Low Expression Filter Threshold
Evaluate model performance across different threshold values and select the best one. 
```{r determine_filter_threshold}
JW26ADS8192 determine_filter_threshold --count ex_counts --meta ex_meta --output ~/JW25ADS8192/tests/testdata/results/
```
Results output a **filtering_analysis.tsv** table in results folder. Success confirmation will read: "Saved as RDS. Done."


#### Step 2: Filter Low Expression Genes
Using the threshold value determined in Step 1, manually replace the `min_count_per_gene` variable and filter. Default `min-ount_per_gene` = 10.
```{r filter_low_exp_genes}
JW26ADS8192 filter_low_exp_genes --count ex_counts --meta ex_meta --output ~/JW25ADS8192/tests/testdata/results/
```
Success confirmation will read: "Saved as RDS. Done." Output **se_filtered.rds** will be save in "~/JW25ADS8192/tests/testdata/results/"


#### Step 3: Differential Gene Expression Analysis via DESeq2
With the remaining filtered genes, preform DESeq2 to analyze the gene expression. Replace `group_var` with the column to be categorized. Default `group_var` = "cell-type". Replace `ref_level` with a specific cell type to be reference. Default `re_level` = "Tconv"
```{r run_DESeq2}
JW26ADS8192 run_DESeq2 --input ~/JW25ADS8192/tests/testdata/results/se_filtered.rds -output ~/JW25ADS8192/tests/testdata/results/  
```
Success confirmation will read: "Saved as RDS. Done." Output **se_dge.rds** will be save in "~/JW25ADS8192/tests/testdata/results/"


#### Step 4: Apply log2_shrinkage on DESeq2 results to improve estimates reliable estimates
Replace `shrinkage` with the appropriate GLM estimator. Default `shrinkage` = "apeglm"
```{r log2_shrinakge}
JWADS8192 log2_shrinkage  --input ~/JW25ADS8192/tests/testdata/results/se_dge.rds -output ~/JW25ADS8192/tests/testdata/results/
```
Success confirmation will read: "Saved as RDS. Done." Results output **dge_shrink.rds** and a **dge_shrink.tsv** table saved in "~/JW25ADS8192/tests/testdata/results/"


#### Step 5: Intrepret the Gene Regulation
Summarize the non-significant and up and down regulated genes in the data set. Replace `p_threshold` with the appropriate adjusted p-value threshold. Default `p_threshold` = 0.05. Replace `fc_threshold` with the appropriate fold-change threshold. Default `fc_threshold` = 0.5. 
```{r gene_regulation_summary}
JW26ADS8192 gene_regulation_summary  --input ~/JW25ADS8192/tests/testdata/results/dge_shrink.rds -output ~/JW25ADS8192/tests/testdata/results/
```
Success confirmation will read: "Saved as RDS. Done." Results output **gene_reg_summary.rds** and a **gene_reg_summary.tsv** table saved in "~/JW25ADS8192/tests/testdata/results/"

#### Step 6: Visualize Expression
Generate a volcano plot to visualize the `gene_reg_summary` results. Use the same `p_threshold` and `fc_threshold` values utilized in Step 5. Replace `set_title` with the correct title. Default `set_title` = "Volcano Plot - Lymph Node Treg vs Tconv". Replace `xlab` with the correct x-axis title. Default `xlab` = "log2 Fold Change (Treg vs Tconv)".
```{r generate_volcano}
JW26ADS8192 generate_volcano --input ~/JW25ADS8192/tests/testdata/results/dge_shrink.rds -output ~/JW25ADS8192/tests/testdata/results/
```
Success confirmation will read: "Saved as RDS. Done." Results output **volcano_plot.rds**, **volcano_plot.pdf**, and **volcano_plot.png** saved in "~/JW25ADS8192/tests/testdata/results/"
```

