---
title: "Differential Expression Analysis via R Studio"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Differential Expression Analysis via R Studio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
### Designed for RNA-seq workflows, **JW26ADS8192** provides a streamlined pipeline to differential expression results. Use a SummarizedExperiment object through the R package interface
```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

---

## Step 1: Install JW26ADS8192 from GitHub
```{r installation}
remotes::install_github("wilsonjewel27/JW26ADS8192")
```

---

## Step 2: Load the following libraries & example dataset
```{r load}
#Required Libraries
library(JW26ADS8192)
library(ggplot2)
library(SummarizedExperiment)
library(DESeq2)
library(apeglm)    
library(stats)
library(rlang)
library(utils)
library(readr)

#example SummarizeExperiment Data Set
data("example_se")

```

---

## Step 3: Start the Differential Gene Expression Analysis

### Step 3.1: Determine the Low Expression Filter Threshold
Evaluate model performance across different threshold values and select the best one.Replace `group_var` with the column to be categorized. Default `group_var` = "cell-type". Replace `ref_level` with a specific cell type to be reference. Default `re_level` = "Tconv". Replace `assay_name` with the appropriate label. Default `assay_name = "counts".  
```{r determine_filter_threshold}
example_se_filtering_assessment <- determine_filter_threshold(
  se_ln             = example_se,
  count_thresholds  = c(0, 1, 5, 10, 20, 50, 100, 200, 500),
  assay_name        = "counts",
  ref_level         = "Tconv",
  group_var         = "cell_type",
  p_threshold       = 0.05
  )

example_se_filtering_assessment

```

### Step 3.2: Filter Low Expression Genes
Using the threshold value determined in Step 1, manually replace the `min_count_per_gene` variable and filter. Use the same SE utilized in Step 1. Default `min_count_per_gene` = 10. Replace `assay_name` with the appropriate label. Default `assay_name = "counts".  
```{r filter_low_exp_genes}
se_filtered <- filter_low_exp_genes(
  se_ln               = example_se,
  min_count_per_group = 10,
  assay_name          = "counts"
  )
```

### Step 3.3: Differential Gene Expression Analysis via DESeq2
With the remaining filtered genes, preform DESeq2 to analyze the gene expression. Replace `group_var` with the column to be categorized. Default `group_var` = "cell-type". Replace `ref_level` with a specific cell type to be reference. Default `re_level` = "Tconv"
```{r run_DESeq}
se_dge <- run_DESeq2(
  se_ln     = se_filtered, 
  group_var = "cell_type", 
  ref_level = "Tconv"
  )
```

### Step 3.4: Apply log2_shrinkage on DESeq2 results to improve estimates.
Replace `shrinkage` with the appropriate GLM estimator. Default `shrinkage` = "apeglm".
```{r log2_shrinkage}
se_dge_shrink <- log2_shrinkage(
  dds       = se_dge, 
  shrinkage = "apeglm"
  )

head(se_dge_shrink)
```

### Step 3.5: Intrepret the Gene Regulation
Summarize the non-significant and up and down regulated genes in the data set. Replace `p_threshold` with the appropriate adjusted p-value threshold. Default `p_threshold` = 0.05. Replace `fc_threshold` with the appropriate fold-change threshold. Default `fc_threshold` = 0.5. 
```{r gene_regulation_summary}
DESeq2_gene_reg_summary <- gene_regulation_summary(
  res_df       = se_dge_shrink,
  p_threshold  = 0.05,
  fc_threshold = 0.5
  )

DESeq2_gene_reg_summary
```

### Step 3.6: Visualize Expression
Generate a volcano plot to visualize the `gene_reg_summary` results. Use the same `p_threshold` and `fc_threshold` values utilized in Step 5. Replace `set_title` with the correct title. Default `set_title` = "Volcano Plot - Lymph Node Treg vs Tconv". Replace `xlab` with the correct x-axis title. Default `xlab` = "log2 Fold Change (Treg vs Tconv)".
```{r generate_volcano}
example_se_volcano <- generate_volcano(
  res_df       = se_dge_shrink,
  fc_threshold = 0.5,
  xlab         = "log2 Fold Change (Treg vs Tconv)",
  set_title    = "Volcano Plot - Lymph Node Treg vs Tconv",
  p_threshold  = 0.05
  )
example_se_volcano
```

### Step 3.7: Export Results
Export the se_dge_shrink, DESeq2_gene_reg_summary, example_se_filtering_assessment data frames as TSV tables and  volcano plots as a PDF and PNG images. Outputs will export as "de_output" to current working directory unless explicitly indicated. 
```r
example_se_exports <- export_outputs(
  res_df         = se_dge_shrink,
  summary_df     = DESeq2_gene_reg_summary,
  filtering_diag = example_se_filtering_assessment,
  volcano        = example_se_volcano,
  output_dir     = file.path(tempdir(), "de_output") 
  )
```
